Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-07-17 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Last Revision: small changes to improve readibility

A two-phase immiscible flow along homogeneous and parallel layers of contrasting petrophysical properties (\ie permeability) is initially simulated. During fluid displacement, crossflow between adjacent layers often occur due to viscous, capillary and/or gravitational forces in discontinuous spatial permeability distribution. Such preferential flows are shown in Figs.~\ref{fem_cv_represent_a}(b-c). In this case, the crossflow dynamics through first and fourth quadrants of the domain is in good qualitative agreement with experiments conducted by \citet{dawe_2008} (see Figs. 5 and 6 of this reference).

@@

\medskip
Viscous finger morphologies were investigated by \citet{guan_2003} based on mathematical formulation and semi-analytic solutions of the coupled Darcy and continuity equations developed by \citet{mclean_1981}. They conducted a sensitivity analysis to investigate fingers' formation, dimensions and branchiness for a range of viscosity ratios $\left(\right.$10$^{2}\le\textit{VR}\le$10$\left.^{4}\right)$ and modified capillary numbers $\left(\right.$632$\le N_{C}^{'}\le$6.32$\times$10$^{7}$, with $N_{C}^{'}= U_{f}\mu\gamma^{-1}(W/b)^{2}$, where $U_{f}$ is the velocity of the finger, $W$ is half Hele-Shaw cell width and $b$ is the cell thickness$\left.\right)$. Finger width $\left(\lambda_{f}\right)$ and volumetric flow rate $\left(\mathcal{Q}\right)$ at the outflow region were correlated through,   
\begin{equation}   
   \mathcal{Q} = U_{f} b \lambda_{f}.\label{eqn:guaneqn} 
\end{equation}  
Here, numerical simulations performed with viscosity ratio of 10 and 150 indicated maximum fingers width of approximately 0.45-0.70 and 0.50-0.90 cm (Fig.~\ref{fig:homoheleshaw_VN10_VN150} a and b), respectively (assuming b = 0.1 cm). This is in close agreement with expected values obtained from Eqn.~\ref{eqn:guaneqn} which indicates maximum finger width ranging from 0.13 to 0.75 (for {\it VR}=10) and from 0.30 to 0.75 ({\it VR}=150, Table~\ref{table:Heleshaw}).

@@

In order to qualitatively investigate the impact of heterogeneity (here represented by changing in the permeability field), numerical simulations were conducted using a prescribed permeability distribution and viscosity ratios of 10 and 150. Absolute permeability ranging from 1.0$\times$10$^{-12}$ to 5.0$\times$10$^{-10}$ cm$^{2}$ (\ie 0.1 $\le$ {\bf K} $\le$ 50 milidarcy) was used in the simulations, which were performed with the same geometry,  mesh resolution, boundary and initial conditions as in previous homogeneous cases(Figs.~\ref{fig:homoheleshaw_VN3}-\ref{fig:homoheleshaw_VN10_VN150}b).

@@

\medskip
Here, two numerical simulations were conducted with fixed and adaptive mesh resolutions to qualitatively assess model capability to capture fingers dynamics. In order to trigger the formation of fingers, regions with sharp permeability gradient were introduced. The computational domain, Fig.~\ref{fig:testcase_heter_domain}a-b, consists of a rectangular geometry fully saturated with a fluid (except by a squared region containing 50$\%$ of a second fluid -- wetting fluid phase, Fig.~\ref{fig:testcase_heter_domain}c) and divided into 5 regions with prescribed synthetic permeability distribution with values between $1$ and $5$. A no-flux boundary condition was imposed across upper and lower borders, whilst pure wetting fluid phase $\left(\text{\ie} S_{1}=1\right)$ is driven into the domain from the left-hand side face at velocity $u_{1}=1$. Viscosity ratio was set to 10.

\medskip
Figures \ref{fig:5regions_fixedmesh} and \ref{fig:5regions_adaptmesh} show fluid displacement in simulations performed with fixed and adaptive meshes. Both simulations started with a mesh of 13068 triangular (\PN[1]{2}) elements, however as flow dynamics evolve the number of elements of the simulation conducted with adaptive mesh oscillates from a minimum of 4400 to a maximum of 16430 (Fig.~\ref{fig:5regions_plottimenodeselements}). In both cases, fingers' development (formation, growth and coalescence) and fluid cross flows (through regions of distinct permeabilities at the top of the domain, $\mathbf{K}=\left\{2,3\right\}$) can be readily noticed. The simulation performed with a fixed and relatively fine mesh was able to capture the progressive development of fingers (as shown in Fig.~\ref{fig:5regions_AdaptZoom}b) prompted by sharp permeability gradient at the interface of regions with $\mathbf{K}=\left\{1,5\right\}$. In regions with no permeability gradient, formation and development of fingers (Fig.~\ref{fig:5regions_AdaptZoom}c) were not captured by the simulation conducted with fixed mesh. In both cases, dynamic adaptive mesh based on perturbation of phase saturations with imposed interpolation error estimate proved to be able to capture the onset of instabilities and fingers' development with relatively little computational overhead, as indicated in Fig.~\ref{fig:5regions_plottimenodeselements}.

@@

\medskip
Simulations were performed in a hexahedron domain of 4 cm $\times$ 2.5 cm $\times$ 10 cm, with permeability distribution of 10$^{-4}\le\mathbf{K}\le$10$^{5}$  milidarcy (Fig.~\ref{fig:3DChannel_PermMesh}). The computational domain was designed to naturally capture preferential flow pathways through channels of large absolute permeability. Pure (wetting) fluid is added into the domain with uniform velocity of 2.5 cm.s$^{-1}$ from the left-hand side face. No-flux boundary conditions are applied to all borders, except at the inlet (left-hand side) and outflow (right-hand side) faces. Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0)=0.80$, which was continuously displaced by the wetting fluid as the simulation progress. The domain was discretised with 235k tetrahedral \PNDG[1]{1} element-pairs.

\medskip
As demonstrated in the preliminary model validation (Section~\ref{section:results_initial_model_validation}), crossflow towards the upper region of the initial part of the domain is due to boundary pressures and diffusivities (both at different vertical layers), \ie caused by large permeability gradients (Fig.~\ref{fig:3DChannel_sat}). As the simulation progresses, non-wetting fluid is driven from regions with lower fluid resistance (central channel), although adjacent region (with low permeability) remains stagnant.

\medskip
Such preferential flow pathways are clearly indicated in Fig.~\ref{fig:3DChannel_planes}, where Darcy velocity vectors overlapped with wetting phase fluid saturation are shown. Fluid displacement occurs mainly in regions of relatively larger permeability, leaving other parts of the domain (\ie regions with low permeability distribution) with little or no momentum, indicating that at these regions the non-wetting fluid (\ie fluid 2) is not effectively displaced.

\medskip
In Sections \ref{section:results_homo_hete}-\ref{section:results_hete_fix_adapt}, finger's formation, growth and branching were triggered by domain heterogeneity represented by permeability differences. The progress of the saturation front and formation and growth of fingers in time and space can be seen in more details in isosurfaces (Fig.~\ref{fig:3DChannel_Isosurf}).

\medskip
A similar numerical simulation (\ie same geometry, initial mesh resolution, initial and boundary conditions) was also performed using dynamically adaptive mesh. During the simulation, mesh resolution ranged from 235k to 643k \PNDG[1]{1} element-pairs (Fig.~\ref{fig:3DChannel_satvel_XYplanes_adapt}a). Distinct preferable flow pathways and fingers are shown in Fig.~\ref{fig:3DChannel_satvel_XYplanes_adapt}b-c and \ref{fig:3DChannel_satvel_XYplanes_adapt}. Saturation of the wetting fluid during the course of the simulation is shown in Fig.~\ref{fig:3DChannel_sat_adapt}. As expected, both simulations (fixed and adaptive mesh, Figs.~\ref{fig:3DChannel_sat} and~\ref{fig:3DChannel_sat_adapt}, respectively) led to similar results, \ie similar flow pathways and fingers development.

\medskip
However, due to low-order accuracy of element-pairs -- \PNDG[1]{1} instead of \PN[1]{2} of previous sections \citep[for full investigation on numerical accuracy associated with these element-pairs, see][]{salinas2015,salinas_2016,salinas_2018,adam_2016, gomes_2017}, and coarser mesh throughout the computational domain, simulation performed with fixed mesh showed larger smeared fluid saturation.

@@

\medskip
Simulations involving formation and growth of viscous fingers were conducted with the CVFEM-based multi-fluid flow simulator Fluidity/IC-FERST model. The model is based upon a dual consistent pressure-velocity representation in CV and FEM spaces and a novel family of FE-pairs, \PN[n]{m} and \PNDG[n]{m}. Saturation and other scalar fields are discretised in CV space and may be fully discontinuous across finite element boundaries. In order to strongly enforce discretised Darcy equations at the boundaries between elements and control volumes, a directional-weighted flux-limited was introduced to take into account discontinuities risen from both, numerical formulation (based on discontinuous-Galerkin finite element method, DGFEM) and control volumes within finite elements.

\medskip
%(2)Summary (1-3 paragraphs) of the main results with the appropriate analysis (\ie how you work has proved the theory);
The numerical formulation was initially validated against laboratory experiments \citep{dawe_2008} to qualitatively assess model functionality to simulate crossflow (\ie preferential flow pathways) in a chequerboard permeability domain. Formation and development of viscous fingers in Hele-Shaw cells (homogeneous domain) were investigated to demonstrate that the onset of interface flow instability occur when VR$\ge$3. Maximum dendritic finger widths obtained from numerical simulations performed with VR=10 and VR=150 largely agree with expected parametric analytic solutions \citep{mclean_1981,guan_2003}. However, when perturbations are imposed into the domain's geophysical properties, \ie spatial variability of absolute permeability, numerical solutions partially agree with analytic solutions (Table~\ref{table:Heleshaw}). This may be due to the domain homogeneity assumption in the analytic solution (Eqns.~\ref {eqn:guaneqn}).

\medskip
Determining the early-onset flow instability is critical to accurately simulate formation, growth and coalescence of fingers, and therefore to: (a) predict displacement (sweep) efficiency and (b) design strategies to either mitigate or avoid fingering and preferential flow pathways. Mesh resolution is thus crucial to investigate Saffmann-Taylor instabilities, and adaptive dynamic mesh technology has proved to be an efficient tool to capture initial interface perturbations. Numerical simulations were performed with VR=10 using fixed and adaptive \PN[1]{2} element-pairs. Results (Figs.~\ref{fig:5regions_plottimenodeselements} and~\ref{fig:5regions_AdaptZoom}) demonstrated that by dynamically adapting the mesh to follow saturation gradients (\ie fluids interfaces), early-onset flow instability and further formation of long multi-fluid fingers, which have themselves split into sub-fingers, can be readily captured (see Figs.~\ref{fig:5regions_fixedmesh}i and~\ref{fig:5regions_adaptmesh}i). Overall computational cost of the simulation conducted with dynamically adapting the mesh were significantly smaller than the simulation conducted with fixed mesh as indicated by the number of elements/nodes (Figs.~\ref{fig:5regions_plottimenodeselements}).

\medskip
Investigation of preferential flow pathways was the focus of Section~\ref{section:results_3D}, where 3D numerical simulations of multi-fluid flow through semi-pervious / impervious channelised domain were performed. Numerical data indicated preferential flow through regions of larger permeability leaving adjacent region (with low permeability) with nearly stagnant fluids (Figs.~\ref{fig:3DChannel_sat}-~\ref{fig:3DChannel_planes}). Order of accuracy of the element-pair and mesh resolution play key-roles to capture directional flow pathways and fingers in 3D simulations.

\medskip
% (3)(optional) 1 paragraph of potential future work. Make sure that (1) and (2) are covered in the present Section.
This work focuses on complexity associated with flow channelling and instability dynamics in heterogeneous porous media. Methods used to simulate such flows are based in a novel high-order CVFEM formulation that accurately preserves sharp fluid saturation gradients associated with contrasting permeability distribution. Future work will include numerical investigation of (a) crossover from capillary fingering (at low fluid velocity) to viscous fingering for immiscible unstable flows, and (b) coupled viscous and density fingering in miscible fluid flows.

>>>>

A two-phase immiscible flow along homogeneous and parallel layers of contrasting petrophysical properties (\ie permeability) is initially simulated. During fluid displacement, crossflow between adjacent layers often occur due to viscous, capillary and/or gravitational forces in discontinuous spatial permeability distribution. Such preferential flows are shown in Figs.~\ref{fem_cv_represent_a}(b-c). In this case, crossflow behaviour from first to fourth quadrants of the domain is in good qualitative agreement with experiments conducted by \citet{dawe_2008} (see Figs. 5-6 of this reference).

@@

\medskip
Viscous finger morphologies were investigated by \citet{guan_2003} based on mathematical formulation and semi-analytic solutions of the coupled Darcy and continuity equations developed by \citet{mclean_1981}. They conducted a sensitivity analysis to investigate fingers' formation, dimensions and branchiness for a range of viscosity ratios $\left(\right.$10$^{2}\le\textit{VR}\le$10$\left.^{4}\right)$ and modified capillary numbers $\left(\right.$632$\le N_{C}^{'}\le$6.32$\times$10$^{7}$, with $N_{C}^{'}= U_{f}\mu\gamma^{-1}(W/b)^{2}$, where $U_{f}$ is the velocity of the finger, $W$ is half Hele-Shaw cell width and $b$ is the cell thickness$\left.\right)$. Finger width $\left(\lambda_{f}\right)$ and volumetric flow rate $\left(\mathcal{Q}\right)$ at the outflow region were correlated through,   
\begin{equation}   
   \mathcal{Q} = U_{f} b \lambda_{f}.\label{eqn:guaneqn} 
\end{equation}  
Here, numerical simulations performed with viscosity ratio of 10 and 150 indicated maximum fingers width of approximately 0.45-0.70 and 0.50-0.90 cm (Fig.~\ref{fig:homoheleshaw_VN10_VN150} a and b), respectively (assuming b = 1 mm). This is in close agreement with expected values obtained from Eqn.~\ref{eqn:guaneqn} which indicates maximum finger width ranging from 0.13 to 0.75 (for {\it VR}=10) and from 0.30 to 0.75 ({\it VR}=150, Table~\ref{table:Heleshaw}).

@@

In order to qualitatively investigate the impact of heterogeneity (here represented by changing in the permeability field), numerical simulations were conducted using a prescribed permeability distribution and viscosity ratios of 10 and 150. Absolute permeability ranging from 1.0$\times$10$^{-12}$ to 5.0$\times$10$^{-10}$ cm$^{2}$ (\ie 0.1 $\le$ {\bf K} $\le$ 50 milidarcy) was used in the simulations, which were performed with the same geometry,  mesh resolution, boundary and initial conditions as in previous homogeneous cases (Figs.~\ref{fig:homoheleshaw_VN3}-\ref{fig:homoheleshaw_VN10_VN150}b).

@@

\medskip
Here, numerical simulations were conducted with fixed and adaptive mesh resolutions to qualitatively assess model capability to capture fingers dynamics. In order to trigger the formation of fingers, regions with sharp permeability gradient were introduced. The computational domain, Fig.~\ref{fig:testcase_heter_domain}a, consists of a rectangular geometry fully saturated with a fluid (except by a squared region containing 50$\%$ of a second fluid -- wetting fluid phase, Fig.~\ref{fig:testcase_heter_domain}c) and divided into 5 regions with prescribed synthetic permeability distribution with values between $1$ and $5$ (Fig.~\ref{fig:testcase_heter_domain}b). A no-flux boundary condition was imposed across upper and lower borders, whilst pure wetting fluid phase $\left(\text{\ie} S_{1}=1\right)$ is driven into the domain from the left-hand side face at velocity $u_{1,x}=1$. Viscosity ratio was set to 10.

\medskip
Figures \ref{fig:5regions_fixedmesh} and \ref{fig:5regions_adaptmesh} show fluid displacement in simulations performed with fixed and adaptive meshes. Both simulations started with a mesh of 13068 triangular (\PN[1]{2}) elements, however as flow dynamics evolve the number of elements of the simulation conducted with adaptive mesh oscillates from a minimum of 4400 to a maximum of 16430 (Fig.~\ref{fig:5regions_plottimenodeselements}). In both cases, fingers' development (formation, growth and coalescence) and fluid cross flows (through regions of sharp permeability gradient at the top of the domain, $\mathbf{K}=\left\{2,3\right\}$) can be readily noticed. The simulation performed with a fixed and relatively fine mesh was able to capture the progressive development of fingers (as shown in Fig.~\ref{fig:5regions_AdaptZoom}b) prompted by sharp permeability gradient at the interface of regions with $\mathbf{K}=\left\{1,5\right\}$. In regions with no permeability gradient, formation and development of fingers (Fig.~\ref{fig:5regions_AdaptZoom}c) were not captured by the simulation conducted with fixed mesh. In both cases, dynamic adaptive mesh based on perturbation of phase saturations with imposed interpolation error estimate proved to be able to capture the onset of instabilities and fingers' development with relatively little computational overhead, as indicated in Fig.~\ref{fig:5regions_plottimenodeselements}.

@@

Simulations were performed in a hexahedron domain of 4 cm $\times$ 2.5 cm $\times$ 10 cm, with permeability distribution of 10$^{-4}\le\mathbf{K}\le$10$^{5}$  milidarcy (Fig.~\ref{fig:3DChannel_PermMesh}). The computational domain was designed to naturally capture preferential flow pathways through channels of large absolute permeability. Pure (wetting) fluid is added into the domain with uniform velocity of 0.5 cm.s$^{-1}$ from the left-hand side face. No-flux boundary conditions are applied to all borders, except at the inlet (left-hand side) and outflow (right-hand side) faces. Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0)=0.80$, which was continuously displaced by the wetting fluid as the simulation progress. The domain was discretised with 235k tetrahedral \PNDG[1]{1} element-pairs.

As demonstrated in the preliminary model validation (Section~\ref{section:results_initial_model_validation}), crossflow towards the upper region of the initial part of the domain is due to boundary pressures and diffusivities (both at different vertical layers), \ie caused by large permeability gradients (Fig.~\ref{fig:3DChannel_sat}). As the simulation progresses, non-wetting fluid is driven from regions with lower fluid resistance (central channel), although adjacent region (with low permeability) remains stagnant.

Such preferential flow pathways are clearly indicated in Fig.~\ref{fig:3DChannel_planes}, where Darcy velocity vectors overlapped with wetting phase fluid saturation are shown. Fluid displacement occurs mainly in regions of relatively larger permeability, leaving other parts of the domain (\ie regions with low permeability distribution) with little or no momentum, indicating that at these regions the non-wetting fluid (\ie fluid 2) is not effectively displaced.

In Sections \ref{section:results_homo_hete}-\ref{section:results_hete_fix_adapt}, finger's formation, growth and branching were triggered by domain heterogeneity represented by permeability differences. The progress of the saturation front and formation and growth of fingers in time and space can be seen in more details in isosurfaces (Fig.~\ref{fig:3DChannel_Isosurf}).

A similar numerical simulation (\ie same geometry, initial mesh resolution, initial and boundary conditions) was also performed using dynamically adaptive mesh. During the simulation, mesh resolution ranged from 235k to 643k \PNDG[1]{1} element-pairs (Fig.~\ref{fig:3DChannel_severalfields_adapt}a). Distinct preferable flow pathways and fingers are shown in Fig.~\ref{fig:3DChannel_severalfields_adapt}-\ref{fig:3DChannel_sat_adapt}. Saturation of the wetting fluid during the course of the simulation is shown in Fig.~\ref{fig:3DChannel_sat_adapt}. As expected, both simulations (fixed and adaptive mesh, Figs.~\ref{fig:3DChannel_sat} and~\ref{fig:3DChannel_sat_adapt}, respectively) led to similar results, \ie similar flow pathways and fingers development.

Low-order accuracy of \PNDG[1]{1} element-pairs (in comparison with \PN[1]{2} used in previous sections) were used in this set of simulations due to the computational overhead \citep[for full investigation on numerical accuracy associated with these element-pairs, see][]{salinas2015,salinas_2016,salinas_2018,adam_2016, gomes_2017}. However, results in this section have shown larger smeared spatial saturation distribution obtained from the fixed mesh simulation indicating numerical diffusion in flows across large permeability gradient.

@@

%\medskip
Simulations involving formation and growth of viscous fingers were conducted with the CVFEM-based multi-fluid flow simulator Fluidity/IC-FERST model. %The model is based upon a dual consistent pressure-velocity representation in CV and FEM spaces and a novel family of FE-pairs, \PN[n]{m} and \PNDG[n]{m}. Saturation and other scalar fields are discretised in CV space and may be fully discontinuous across finite element boundaries. In order to strongly enforce discretised Darcy equations at the boundaries between elements and control volumes, a directional-weighted flux-limited was introduced to take into account discontinuities risen from both, numerical formulation (based on discontinuous-Galerkin finite element method, DGFEM) and control volumes within finite elements.
%\medskip
%(2)Summary (1-3 paragraphs) of the main results with the appropriate analysis (\ie how you work has proved the theory);
The numerical formulation was initially validated against laboratory experiments \citep{dawe_2008} to qualitatively assess model functionality to simulate crossflow (\ie preferential flow pathways) in a chequerboard permeability domain. Formation and development of viscous fingers in Hele-Shaw cells (homogeneous domain) were investigated to demonstrate that the onset of interface flow instability occur when VR$\ge$3. Maximum dendritic finger widths obtained from numerical simulations performed with VR=10 and VR=150 largely agree with expected parametric analytic solutions \citep{mclean_1981,guan_2003}. However, when perturbations are imposed into the domain's geophysical properties, \ie spatial variability of absolute permeability, numerical solutions partially agree with analytic solutions (Table~\ref{table:Heleshaw}). This may be due to the domain homogeneity assumption in the analytic solution (Eqns.~\ref {eqn:guaneqn}).

%\medskip
Determining the early-onset flow instability is critical to accurately simulate formation, growth and coalescence of fingers, and therefore to: (a) predict displacement (sweep) efficiency and (b) design strategies to either mitigate or avoid fingering and preferential flow pathways. Mesh resolution is thus crucial to investigate Saffmann-Taylor instabilities, and adaptive dynamic mesh technology has proved to be an efficient tool to capture initial interface perturbations. Numerical simulations were performed with VR=10 using fixed and adaptive \PN[1]{2} element-pairs. Results (Figs.~\ref{fig:5regions_plottimenodeselements} and~\ref{fig:5regions_AdaptZoom}) demonstrated that by dynamically adapting the mesh to follow saturation gradients (\ie fluids interfaces), early-onset flow instability and further formation of long multi-fluid fingers, which have themselves split into sub-fingers, can be readily captured (see Figs.~\ref{fig:5regions_fixedmesh}i and~\ref{fig:5regions_adaptmesh}i). Overall computational cost of the simulation conducted with dynamically adapting the mesh were significantly smaller than the simulation conducted with fixed mesh as indicated by the number of elements/nodes (Figs.~\ref{fig:5regions_plottimenodeselements}).

%\medskip
Investigation of preferential flow pathways was the focus of Section~\ref{section:results_3D}, where 3D numerical simulations of multi-fluid flow through semi-pervious / impervious channelised domain were performed. Numerical data indicated preferential flow through regions of larger permeability leaving adjacent region (with low permeability) with nearly stagnant fluids (Figs.~\ref{fig:3DChannel_sat}-~\ref{fig:3DChannel_planes}). Order of accuracy of the element-pair and mesh resolution play key-roles to capture directional flow pathways and fingers in 3D simulations.

%\medskip
% (3)(optional) 1 paragraph of potential future work. Make sure that (1) and (2) are covered in the present Section.
This work focuses on complexity associated with flow channelling and instability dynamics in heterogeneous porous media. Methods used to simulate such flows are based in a novel high-order CVFEM formulation that accurately preserves sharp fluid saturation gradients associated with contrasting permeability distribution. Future work will include numerical investigation of (a) crossover from capillary fingering (at low fluid velocity) to viscous fingering for immiscible unstable flows, and (b) coupled viscous and density fingering in miscible fluid flows.

### 2018-07-17 Papers/ViscousInstabilities/article_figure1.tex
Last Revision: small changes to improve readibility

%%%%
%%%%  FIGURE 
%%%%
\begin{figure}[h]
  \hbox{\hspace{2cm}
  \includegraphics[width=0.75\textwidth]{./Pics1/Saffman_homogeneous/MeshIndependence.pdf}}
  \caption{Mesh convergence analysis for Hele-Shaw cells flow simulations: non-wetting phase saturation profiles through a center-line across the domain (diagonal) with grid resolution ranging from 1881 to 3546 \PN[1]{2} triangular elements.}
\label{fig:MeshDependenceAnalysis}
\end{figure}
\clearpage

@@

%%%%
%%%%  FIGURE
%%%%
\begin{figure}[ht]
\vbox{
\hbox{\includegraphics[width=.5\textwidth]{./Pics1/Saffman_homogeneous_VR10/ST_Homog_VR10_D2201_W2b.pdf}
       \includegraphics[width=.5\textwidth]{./Pics1/Saffman_homogeneous_VR150/ST_Homog_VR150_D5003_W2b.pdf}}
\hbox{\hspace{0.25cm} (a) {\it VR}=10 (homogeneous) \hspace{1.5cm} (b) {\it VR}=150 (homogeneous) }
\vspace{0.5cm}
\hbox{\includegraphics[width=.5\textwidth]{./Pics1/Saffman_heterogeneous_VR10/ST_Heterog_VR10_D11000_W2b.pdf}
       \includegraphics[width=.5\textwidth]{./Pics1/Saffman_heterogeneous_VR150/ST_Heterog_VR150_D6500_W2b.pdf}}
\hbox{\hspace{0.25cm} (c) {\it VR}=10 (heterogeneous) \hspace{1.5cm} (d) {\it VR}=150 (heterogeneous) }
}
\caption{Isosurfaces of simulated flows in Hele-Shaw cells with viscosity ratios of 10 (a and c) and 150 (b and d). Top and bottom rows describe simulations performed with constant \textit{i.e. homogeneous with \textbf{K}}= $10^{-10}$ \textit{${cm}^{2}$} and randomly distributed (i.e. heterogeneous, Fig.~\ref{fig:HeleShawHeter_VR10}a) permeabilities. Width of largest fingers for homogeneous cases are approximately $0.70$ and $0.90$cm (\textit{VR}=$10$ and \textit{VR}=$150$, respectively), whereas for heterogeneous cases are $0.56$ and $0.40$cm. Results for homogeneous cases are in good agreement with values obtained from \citet{guan_2003}'s analytic solution.}
\label{fig:homoheleshaw_VN10_VN150}
\end{figure}
\clearpage

@@

%%%
%%% FIGURE XXXXXX
%%%
  \begin{figure}[ht]
  \vbox{\vspace{-1.cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D0b.pdf}
            \includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D250b.pdf}} 
            %\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D350b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(a) t=0.00 \hspace{4cm} (f) t= 0.35}\vspace{-.1cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D100b.pdf} 
            \includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D700b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(b) t=0.10 \hspace{4cm} (g) t= 0.70}\vspace{-.1cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D150b.pdf} 
            \includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D1000b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(c) t=0.15 \hspace{4cm} (h) t= 1.00}\vspace{-.1cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D200b.pdf} 
            \includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D1500b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(d) t=0.20 \hspace{4cm} (i) t= 1.50}\vspace{-.1cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D350b.pdf}}
      %\hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D250b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(e) t=0.25}\vspace{-.1cm}}
\caption{Impact of mesh resolution on capturing flow instabilities: snapshots of saturation field through 1.50 unit-time of numerical simulation performed with fixed mesh (of 13068 elements) and {\it VR}=10.}
\label{fig:5regions_fixedmesh}
\end{figure}
\clearpage

@@

%%%
%%% FIGURE XXXXXX
%%%
\begin{figure}[ht]
\vbox{\vspace{-.5cm}
      \hbox{\includegraphics[width=\textwidth]{./Pics1/3D_Channel/Test_SlowNew_MeshPermeability.pdf} }}
\caption{3D channel flow: Permeability distribution within the computational domain, containing 235242 \PNDG[1]{1} element-pairs.}
\label{fig:3DChannel_PermMesh}
\end{figure}
\clearpage

>>>>

%%%%
%%%%  FIGURE 
%%%%
\begin{figure}[h]
  \hbox{\hspace{2cm}
  \includegraphics[width=0.75\textwidth]{./Pics1/Saffman_homogeneous/MeshIndependence.pdf}}
  \caption{Mesh convergence analysis for Hele-Shaw cells flow simulations (VR=1): non-wetting phase saturation profiles through a center-line across the domain (diagonal) with grid resolution ranging from 1881 to 3546 \PN[1]{2} triangular elements. }
\label{fig:MeshDependenceAnalysis}
\end{figure}
\clearpage

@@

%%%%
%%%%  FIGURE
%%%%
\begin{figure}[ht]
\vbox{
\hbox{\includegraphics[width=.5\textwidth]{./Pics1/Saffman_homogeneous_VR10/ST_Homog_VR10_D2201_W2b.pdf}
       \includegraphics[width=.5\textwidth]{./Pics1/Saffman_homogeneous_VR150/ST_Homog_VR150_D5003_W2b.pdf}}
\hbox{\hspace{0.25cm} (a) {\it VR}=10 (homogeneous) \hspace{1.5cm} (b) {\it VR}=150 (homogeneous) }
\vspace{0.5cm}
\hbox{\includegraphics[width=.5\textwidth]{./Pics1/Saffman_heterogeneous_VR10/ST_Heterog_VR10_D11000_W2b.pdf}
       \includegraphics[width=.5\textwidth]{./Pics1/Saffman_heterogeneous_VR150/ST_Heterog_VR150_D6500_W2b.pdf}}
\hbox{\hspace{0.25cm} (c) {\it VR}=10 (heterogeneous) \hspace{1.5cm} (d) {\it VR}=150 (heterogeneous) }
}
\caption{Isosurfaces of simulated flows in Hele-Shaw cells with viscosity ratios of 10 (a and c) and 150 (b and d). Top and bottom rows describe simulations performed with constant $\left(\text{\ie homogeneous with }\textbf{K}\text{ = 10}^{-10}\text{ cm}^{2}\right)$ and randomly distributed (\ie heterogeneous, Fig.~\ref{fig:HeleShawHeter_VR10}a) permeabilities. Width of largest fingers for homogeneous cases are approximately $0.70$ and $0.90$cm (\textit{VR}=$10$ and \textit{VR}=$150$, respectively), whereas for heterogeneous cases are $0.56$ and $0.40$cm. Results for homogeneous cases are in good agreement with values obtained from \citet{guan_2003}'s analytic solution.}
\label{fig:homoheleshaw_VN10_VN150}
\end{figure}
\clearpage

@@

%%%
%%% FIGURE XXXXXX
%%%
  \begin{figure}[ht]
  \vbox{\vspace{-1.cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D0b.pdf}
            \includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D250b.pdf}} 
            %\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D350b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(a) t=0.00 \hspace{4cm} (f) t= 0.35}\vspace{-.1cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D100b.pdf} 
            \includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D700b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(b) t=0.10 \hspace{4cm} (g) t= 0.70}\vspace{-.1cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D150b.pdf} 
            \includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D1000b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(c) t=0.15 \hspace{4cm} (h) t= 1.00}\vspace{-.1cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D200b.pdf} 
            \includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D1500b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(d) t=0.20 \hspace{4cm} (i) t= 1.50}\vspace{-.1cm}
      \hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D350b.pdf}}
      %\hbox{\includegraphics[width=.45\textwidth]{./Pics1/Section4_4/5r_po_left_inlet_D250b.pdf}}
      \vspace{-.1cm}\hbox{\hspace{2.cm}(e) t=0.25}\vspace{-.1cm}}
\caption{Impact of mesh resolution on capturing flow instabilities: snapshots of saturation field through 1.50 unit-time of numerical simulation performed with fixed mesh (i) of 13068 elements and {\it VR}=10.}
\label{fig:5regions_fixedmesh}
\end{figure}
\clearpage

@@

%%%
%%% FIGURE XXXXXX
%%%
\begin{figure}[ht]
\vbox{\vspace{-.5cm}
      \hbox{\includegraphics[width=\textwidth]{./Pics1/3D_Channel/Test_SlowNew_MeshPermeability.pdf} }}
\caption{3D channel flow: Permeability distribution within the computational domain, containing 235242 \PNDG[1]{1} tetrahedra element-pairs.}
\label{fig:3DChannel_PermMesh}
\end{figure}
\clearpage
